BIP-Predictor

### 2025-06-04 Sweave/BIP_predictor.Rnw
up

# Load the required R-libraries
# Standard filter package
library(mFilter)
# Multivariate time series: VARMA model for macro indicators: used here for simulation purposes only
library(MTS)
# HAC estimate of standard deviations in the presence of autocorrelation and heteroscedasticity
library(sandwich)
# Library for Diebold-Mariano test of equal forecast performance
library(multDM)
# GARCH model: for improving regression estimates
library(fGarch)
# Library for tables
library(xtable)
# Packages for Ridge and LASSO
library(MASS)
library(glmnet) 
# Extended time series
library(xts)
# Extended time series
library(pROC)

# Load the relevant M-SSA functionalities
# M-SSA functions
source(paste(getwd(),"/R/functions_MSSA.r",sep=""))
# Load signal extraction functions used for JBCY paper (relies on mFilter)
source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
# Utility functions for M-SSA, see tutorial 
source(paste(getwd(),"/R/M_SSA_utility_functions.r",sep=""))
source(paste(getwd(),"/R/M_SSA_paper_functions.r",sep=""))
# ROC curve: plots, AUC
source(paste(getwd(),"/R/ROCplots.r",sep=""))

@@

# Settings for M-SSA
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
f_excess<-c(5,rep(4,length(select_vec_multi)-1))
# In-smaple span for VAR
date_to_fit<-"2008"
# We can select type of VAR model
VAR_type="BVAR"
VAR_type="VAR"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 1.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}

# Filter length
L<-31
# Publication lag: one quarter for BIP
lag_vec<-c(1,rep(0,ncol(x_mat)-1))

@@

    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj,file=paste(path.result,"/BVAR",sep=""))
  }
# Without Pandemic  
  data_fit<-x_mat_wc
  if (VAR_type=="VAR")
  {

    set.seed(12)
    V_obj<-VARMA(data_fit,p=p,q=q)
# Apply regularization: see vignette MTS package
    threshold<-1.5
    V_obj_wc<-refVARMA(V_obj, thres = threshold)
    save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
  }
  if (VAR_type=="BVAR")
  {
    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
    #   low values <=> low precision
    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj_wc <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj_wc,file=paste(path.result,"/BVAR_wc",sep=""))
  }

} else
{
  if (VAR_type=="VAR")
  {
    load(file=paste(path.result,"/VAR",sep=""))
    load(file=paste(path.result,"/VAR_wc",sep=""))
  } else
  {
    load(file=paste(path.result,"/BVAR",sep=""))
    load(file=paste(path.result,"/BVAR_wc",sep=""))
  }
}

@@

# We rely on data with Pandemic
#   -M-SSA is not affected since the in-sample span ends in 2008
if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)

@@

  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
  if (VAR_type=="VAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/nowcast_obj",sep=""))
    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/BV_nowcast_obj",sep=""))
    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
  }
} else
{
  if (VAR_type=="VAR")
  {
    load(file=paste(path.result,"/nowcast_obj",sep=""))
    load(file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    load(file=paste(path.result,"/BV_nowcast_obj",sep=""))
    load(file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
  }

@@

if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
# Without pandemic  
  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
  if (VAR_type=="VAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/mssa_indicator_obj",sep=""))
    save(mssa_indicator_obj_wc,file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/BV_mssa_indicator_obj",sep=""))
    save(mssa_indicator_obj_wc,file=paste(path.result,"/BV_mssa_indicator_obj_wc",sep=""))
  }

@@

# Run the M-SSA wrapper, see tutorial 7.2
#   -The function computes M-SSA and M-MSE for each forecast horizon h in h_vec
if (recompute_results)
{
  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)

>>>>

# Load the required R-libraries
# Standard filter package
library(mFilter)
# Multivariate time series: VARMA model for macro indicators: used here for simulation purposes only
library(MTS)
# HAC estimate of standard deviations in the presence of autocorrelation and heteroscedasticity
library(sandwich)
# Library for Diebold-Mariano test of equal forecast performance
library(multDM)
# GARCH model: for improving regression estimates
library(fGarch)
# Library for tables
library(xtable)
# Packages for Ridge and LASSO
library(MASS)
library(glmnet) 
# Extended time series
library(xts)
# Extended time series
library(pROC)
# Elastic net for estimating VAR
library(elasticnet)

# Load the relevant M-SSA functionalities
# M-SSA functions
source(paste(getwd(),"/R/functions_MSSA.r",sep=""))
# Load signal extraction functions used for JBCY paper (relies on mFilter)
source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
# Utility functions for M-SSA, see tutorial 
source(paste(getwd(),"/R/M_SSA_utility_functions.r",sep=""))
source(paste(getwd(),"/R/M_SSA_paper_functions.r",sep=""))
# ROC curve: plots, AUC
source(paste(getwd(),"/R/ROCplots.r",sep=""))
# Elastic net functions for VAR estimation
source(paste(getwd(),"/R/elastic_net_functions.r",sep=""))

@@

# Settings for M-SSA
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
f_excess<-c(5,rep(4,length(select_vec_multi)-1))
# In-smaple span for VAR
date_to_fit<-"2008"
# Filter length
L<-31

# We can select type of VAR model
VAR_type="BVAR"
VAR_type="VAR"
VAR_type="elastic_net"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 1.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}
if (VAR_type=="elastic_net")
{
# Marc's settings for VAR  
  p<-2
  q<-0
# Length of MA inversion (impulse response function)  
  nirf<-L
# Number of non-zero coefficients in VAR for each dimension  
  nsteps<-5
}

# Publication lag: one quarter for BIP
lag_vec<-c(1,rep(0,ncol(x_mat)-1))

@@

    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj,file=paste(path.result,"/BVAR",sep=""))
  }
  if (VAR_type=="elastic_net")
  {
    V_obj<-enetVAR(data_fit, p, nIRF, nsteps  )

    save(V_obj,file=paste(path.result,"/en_VAR",sep=""))
  }

# Without Pandemic  
  data_fit<-x_mat_wc
  if (VAR_type=="VAR")
  {

    set.seed(12)
    V_obj<-VARMA(data_fit,p=p,q=q)
# Apply regularization: see vignette MTS package
    threshold<-1.5
    V_obj_wc<-refVARMA(V_obj, thres = threshold)
    save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
  }
  if (VAR_type=="BVAR")
  {
    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
    #   low values <=> low precision
    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj_wc <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj_wc,file=paste(path.result,"/BVAR_wc",sep=""))
  } 
  if (VAR_type=="elastic_net")
  {
    V_obj_wc<-enetVAR(data_fit, p, nIRF, nsteps  )

    save(V_obj_wc,file=paste(path.result,"/en_VAR_wc",sep=""))
  }

} else
{
  if (VAR_type=="VAR")
  {
    load(file=paste(path.result,"/VAR",sep=""))
    load(file=paste(path.result,"/VAR_wc",sep=""))
  } 
  if (VAR_type=="BVAR")
  {
    load(file=paste(path.result,"/BVAR",sep=""))
    load(file=paste(path.result,"/BVAR_wc",sep=""))
  }
  if (VAR_type=="elastic_net")
  {
    load(file=paste(path.result,"/en_VAR",sep=""))
    load(file=paste(path.result,"/en_VAR_wc",sep=""))
  }
}

@@

# We rely on data with Pandemic
#   -M-SSA is not affected since the in-sample span ends in 2008
if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)

@@

  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
  if (VAR_type=="VAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/nowcast_obj",sep=""))
    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/BV_nowcast_obj",sep=""))
    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
  }
} else
{
  if (VAR_type=="VAR")
  {
    load(file=paste(path.result,"/nowcast_obj",sep=""))
    load(file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    load(file=paste(path.result,"/BV_nowcast_obj",sep=""))
    load(file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
  }

@@

if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
# Without pandemic  
  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
  if (VAR_type=="VAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/mssa_indicator_obj",sep=""))
    save(mssa_indicator_obj_wc,file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
  }
  if (VAR_type=="BVAR")
  {
    save(mssa_indicator_obj,file=paste(path.result,"/BV_mssa_indicator_obj",sep=""))
    save(mssa_indicator_obj_wc,file=paste(path.result,"/BV_mssa_indicator_obj_wc",sep=""))
  }

@@

# Run the M-SSA wrapper, see tutorial 7.2
#   -The function computes M-SSA and M-MSE for each forecast horizon h in h_vec
if (recompute_results)
{
  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
